%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% abstrac-en.tex
%% NOVA thesis document file
%%
%% Abstract in English
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Recent times have seen a very big increase of applications being moved to the cloud. The ease of use and deployment, unlimited resources and the reduced costs have convince all kinds of corporations, start-ups and enterprises to move their infrastructure, applications and services to a cloud provider. The increase flow of data over the web, as well of the increase of data stored and processed off site has reach the knowledge of hackers and so, it has raised big concerns over the security researchers community. In order to keep all this data secure, hardware and security companies have been working hard to provide technologies and solutions based on hardware components that would guarantee applications to run in a secure environment from any external or internal attacks.

The recent advent of hardware-based trusted execution environments provides isolated execution even from an untrusted operating system and possible hardware-based attacks. As the processor provides such “shielded” execution environments, their use will allow cloud users to run applications securely, for example on the remote cloud servers, whose operating systems and hardware are exposed to potentially malicious remote attackers and non-controlled system administrators’ staff.

This dissertation will design, implement and evaluate experimentally a Trusted and Privacy-Enhanced In Memory Data Structure Store. The proposal will be designed to run on a trusted execution environment supported by Intel SGX technology, and will provide an unmodified and completely abstract API to the user despite all the additional security features. It will focus on implementing a storage system that will be kept encrypted in computer memory for all the data lifecycle and guarantee that operations that have critical security issues will be performed isolated and safely.

This dissertation proposes a comparison between a multiple levels of secure replication to provide an available, reliable and dependable system thus offering high availability for data store access, built-in replication, a LRU eviction model, support for transactions and options for on-disk persistence.

% Palavras-chave do resumo em Inglês
\begin{keywords}
Hardware Security; Privacy-Enhanced Data Store; Isolated Environments; Trusted Computing; Availability; Reliability.
\end{keywords} 
