%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgements.tex
%% NOVA thesis document file
%%
%% Text with acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgements

I would like to thank the FCT university and all its faculty, especially professor Henrique João Lopes Domingos who helped me all throughout this dissertation and always took the time to help and motivate me. 

I would also take this opportunity to thank my family, as they have supported me unconditionally not only during the college years but all my life. I thank Sofia Miranda who tolerated my mood swings and my stress during this period, calmed me down and has listen to my outbursts and thesis talk even though computer science is not her area.

I would like to thank the whole WATERDOG team, they took me in since my graduation and helped me start my adult career making me feel like one of their own since the beginning and with an  especial thank you to Bruno Félix and Pedro Coutinho who picked up the slack of the days I've missed work to have time for this dissertation. They are extraordinary and amazing work colleagues.

Lastly, I want to thank my friends who went through the dissertation at the same time, especially João Reis and Tiago Sousa, we "suffered" together but now it is done. Also to my friends who already overcame this objective or are still to do it. You know who you are.

\vspace{20mm}

Thanks to all, could not have done it without you...